Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-24 main.Rmd
add DOIs to references

Time deconstructions can also be based on the hierarchical structure of time. For example, hours are nested within days, days within weeks, weeks within months, and so on. Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour. The lubridate package [@G_Grolemund2011-vm] provides tools to access and manipulate common date-time objects. But most of its accessor functions are limited to single-order-up granularities.

@@

A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.  <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn). Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity. A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson2000-mr.

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and quasi-circular granularities.

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf]. For more variations of calendar algebra operations, see @Ning2002-tf.

@@

```{r write-bib, eval = FALSE}
write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas"), file = "rpkgs.bib")
```

>>>>

Time deconstructions can also be based on the hierarchical structure of time. For example, hours are nested within days, days within weeks, weeks within months, and so on. Hence, it is possible to construct single-order-up granularities such as second of the minute, or multiple-order-up granularities such as second of the hour. The lubridate package [@Grolemund2011-vm] provides tools to access and manipulate common date-time objects. But most of its accessor functions are limited to single-order-up granularities.

@@

A discrete time model often uses a fixed smallest linear granularity named by @Bettini1998-ed **bottom granularity**.  <!--so that the the underlying time domain can be described as a sequence of non-decomposable, consecutive time intervals of some fixed, minimal duration called _chronons_ (@Jensen1998-qn). Chronons may be grouped into granules of a bottom granularity and --> Granules in bottom granularity or any finer granularity may be aggregated in some manner to form larger granules belonging to a coarser granularity. A system of multiple granularities in lattice structures is referred to as a **calendar** by @Dyreson_2000.

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning_2002]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom one. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and quasi-circular granularities.

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days. We can consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning_2002]. For more variations of calendar algebra operations, see @Ning_2002.

@@

```{r write-bib, eval = FALSE}
write_bib(c("dplyr", "tidyr", "tsibble", "magrittr", "rlang", "gravitas", "knitr", "rmarkdown"), file = "rpkgs.bib")
```
